Unsupervised Detection and Tracking of Multiple Objects with Dependent Dirichlet Process Mixtures

### 2011-12-18 ltx/paper.tex
more edits. removed pretty worm picture

We define the task of automated detection and tracking of arbitrary objects in videos to be the fulfillment of three actions: determining the spatiotemporal regions of a video that constitute objects (`extraction'), finding the positions and/or shapes of distinct objects (`localization'), and maintaining the identities of the detected objects over time (`tracking'). An ability to carry out these tasks in an automated manner is useful for many fields that make use of video data, including robotics, video surveillance, computational biology, and video summarization. Algorithms proving to be efficacious for a variety of objects and video types are particularly applicable. This paper serves to introduce a new technique involving the use of dependent Dirichlet processes mixture models, which we hope will provide the foundation for a class of unsupervised algorithms that can detect and track arbitrary objects in a wide range of videos.

Research in the past ten years involving the development of methods for general detection and tracking of objects has tended to focus on one of either extraction, localization, or tracking; attempts to integrate all three tasks and produce a system aimed at multiple arbitrary objects and diverse video types was not often a primary focus, or was attempted in an ad-hoc manner to test or apply a more specific technique. Research aimed at developing methods to accomplish the three tasks in a cohesive manner has been pursued more recently to produce algorithms that can, in a fully automated way, accomplish unsupervised detection and tracking--or in short, that can determine the spatiotemporal region of a video occupied by each distinct object. A widely applied, though unsophisticated, example of general detection and tracking is the technique of blob tracking. Blob tracking typically refers to methods that perform extraction and localization of objects at each frame independently--for example, by assuming that pixels representing an object are distinguishable from pixels representing the background via color, motion, or texture information, and segmenting the extracted data at each frame into `blobs' to accomplish localization. A number of tracking techniques based on finding blobs in consecutive frames of a video with similar locations and appearances have been attempted to maintain the identity of detected blobs over time. The main challenges, however, associated with blob tracking involve the difficulty of segmenting neighboring or occluding objects in a given frame \cite{zhao2004tracking}, preventing multiple tracked blobs from collapsing onto and tracking a single object \cite{vermaak_2003}. These difficulties stem from the typical blob filter's need to perform detection via segmentation on individual frames and lack of a principled way to perform tracking while resisting the collapse of independent tracks.

This paper presents the use of a Bayesian nonparametric mixture model for unsupervised multiple object detection and tracking. This method is intended to allow for robust detection of arbitrary objects (particularly in cases where frame-by-frame segmentation is difficult) and better isolation of independent objects during tracking. Specifically, this work describes how a Dirichlet process mixture model can accomplish these goals when it is applied to data that may be extracted in a general manner from videos. We will demonstrate how this generative model may be applied to types of noisy data that can be quickly extracted from objects in videos, and how this model allows for automated detection and tracking. The model is first presented in an abstract form, to allow for it to be used with arbitrary extraction procedures and object appearance representations that may be desired in future studies. We then formulate a specific model for data gained during a basic extraction procedure and describe inference algorithms that allow object localization and tracking to be carried out. We demonstrate our implementation on a number of synthetic and benchmark datasets and describe and compute performance metrics to quantify results.

@@

The task of extraction is related to the tasks of foregound segmentation, background modeling, motion detection, and appearance feature extraction. In general, these methods aim to extract features from regions of a video frame that are considered objects or are distinguished in some way from the background of a scene. Areas of research in this field include attempts to handle situations where the background displays motion or other dynamic characteristics, reduce noise caused by the false extraction of backgrounds, handle abberant image effects present in some videos, and operate accurately on scenes involving illumination changes. These procedures can usually be classified as either model-based methods or heuristic-based methods. Model-based methods emphasize the detection of shapes of regions containing objects accurately and reducing background noise. For example, foreground extraction has been robustly carried out by modeling each pixel over time in order to infer a probability distribution over pixel values; if the pixel takes on a very unlikely value at a future time, it is marked as a foreground pixel at that point \citep{stauffer_1999, elgammal_2000, elgammal_2002}. Note that these techniques treat each pixel independently; this has a tendency to cause errors in foreground detection, resulting in fragmentation of foreground images (which is problematic as many of these methods intend to blob the foreground objects after detection in order to perform segmentation and locate objects' centroids). Heuristic-based methods perform extracting by looking for temporal changes in pixel value characteristics. These methods are often less accurate than model-based methods, but also far less computationally expensive. Frame differencing and background subtraction are two such methods, and involve comparing pixel values, or groups of pixel values, between two images, to detect a change surpassing some threshold. The goal of these techniques is to find locations in a scene that display motion, and to deem these moving areas the foreground of a scene.  Often, background subtraction refers to comparing an image containing targets with an image of the background without targets or with some model of the background that is learned as the video progresses (in these cases, objects that do not move throughout a scene will be treated as the background), while frame or image differencing refers to comparing two subsequent images in a video. Frame differencing has been used as the sole extraction method for object localization or tracking schemes with success \citep{pece_2002, beleznai_2006, chu_2007}, and also as a secondary data extraction method to help improve the accuracy of object tracking schemes \citep{perez_2002}. In addition to the two techniques mentioned above, there exist other ways of extracting movement from an image, such as through techniques involving optical flow (a calculation based on the `generalized gradient model' of an image, which attempts to capture the speed and direction of movement over areas in the image) \citep{horn_1981, bobick_2001}. In \citep{black_2000}, a motion detection and object tracking method is developed based on the relative motion of moving objects' boundaries, which are again found through analysis of optical flow.

The task of localization is closely related to the topics of object detection, position-finding, segmentation, and target recognition. These methods are mainly concerned with finding the positions and/or shapes of objects in a given video frame, or with discerning between objects similar in position or appearance. There has been a focus in this field on developing techniques for representing an object's appearance, both to allow for localization of objects on a per-frame basis and to allow for consistent localization of a given object over time. Research into the unsupervised detection of arbitrary objects (which we define as methods that accomplish both extraction and localization) has included work involving image segmentation, parsing, and pixel clustering \willie{must cite here}. These topics usually relate to the automated partitioning of images into meaningful regions. Note that the popular phrase `object detection' is also typically used to refer to procedures that carry out extraction and localization; however, it is often reserved to denote methods that are tailored to find the position of specified objects or object types (as opposed to arbitrary object detection), or to denote supervised methods, where extraction is peformed implicitely during searches for specified objects and localization procedures.

The task of tracking incorporates the topics of video tracking, data association, filtering, point matching, and `tracking of non-rigid objects'. In general, these tasks aim to maintain the localization of the position and/or shape of an object over time. Areas of interest in this field involve developing metrics for judging the similarity between two object representations (for example, potential representations of the same object at adjacent frames in a video), techniques that provide a way to conduct efficient searches over a frame for regions that are similar to a given object, and methods for predicting the future position or appearance of an object.

@@

\subsection{Data}
\label{sec:data}
%
We defined the task of extraction to be the unsupervised determination of which spatiotemporal regions of a video constitute objects, and outlined a number of extraction methods in Section~\ref{sec:relatedwork}. The features of the data are characteristics of pixels in the regions of the video that contain objects, and the data are gained during extraction. For example, at its most basic, an extraction dataset is comprised of the three-dimensional points $\bold{x} = ( x_{1}, x_{2}, t ) \in \mathbb{R}^{2} \times \mathbb{Z}_{+}$, where $x_{1}$ and $x_{2}$ are the two spatial dimensions of a pixel contained in a region representing an object in a given frame, and $t$ is the discrete time index of the frame. These three features are always available once regions containing objects are discerned, and are necessary features for modeling (as position information is key for reasonable tracking and time information is necessary for maintaining the temporal dependencies in the data).

@@

We will use the phrase `baseline' three-dimensional extraction data to refer to the $\bold{x} = ( x_{1}, x_{2}, t ) \in \mathbb{R}^{2} \times \mathbb{Z}_{+}$ from pixels present in regions containing objects. These data are distributed roughly as worm-like shapes that follow the paths of individual objects as they move, when they are plotted in three dimensions (where we let the frame number $t$, our discrete representation of time, ascend the vertical axis and the spatial positions occupy the horizontal axes). An example of data with such a distribution can been seen in Figure~\ref{fig:baselinedata}. The data consists of the three-dimensional baseline features and was gained by extraction peformed on a 50 frame sequence showing five ants moving across the video scene.
% \begin{figure*}[h]
%         \center{\includegraphics[width=85mm]{../img/usbh_data_2.png}}
%         \caption{\label{fig:baselinedata} Baseline three-dimensional extraction data gained from a video segment containing 5 moving ants. The vertical axis denotes the frame number and the horizontal axes denote $d_{1}$ and $d_{2}$, the spatial coordinates of pixels within a given frame. Markers are colored by position for ease of viewing. \willie{show a number of baseline data here. consider whether colors help view or just add confusion.}}
% \end{figure*}

\begin{figure}[h]
  \centering
  \subfloat[]{\label{fig:baselinedata1}\includegraphics[width=0.33\textwidth]{../img/pets2001_data_1.png}} 
  \subfloat[]{\label{fig:baselinedata1}\includegraphics[width=0.33\textwidth]{../img/usbh_data_3.png}} 
  \subfloat[]{\label{fig:baselinedata2}\includegraphics[width=0.33\textwidth]{../img/traffic2_data_1.png}}\\
  \subfloat[]{\label{fig:baselinedata1}\includegraphics[width=0.33\textwidth]{../img/pets2009_data_1.png}} 
  \subfloat[]{\label{fig:baselinedata1}\includegraphics[width=0.33\textwidth]{../img/pets2009_data_2.png}} 
  \subfloat[]{\label{fig:baselinedata2}\includegraphics[width=0.33\textwidth]{../img/pets2009_data_3.png}}
  \caption{Baseline extracted features from the (a.) PETS2001 Dataset, (b.) Ants Dataset, (c.) Traffic Dataset, and PETS2009/2010 benchmark dataset, frames (a.) 1-200, (b.) 201-400, (c.) 401-600. Time is shown on the vertical axis. Marker color corresponds with spatial coordinates and is shown for ease of viewing.}
  \label{fig:baselinedata_both}
\end{figure}
Inference performed on the time dependent Dirichlet process mixture model (Section~\ref{sec:inference}) carries out clustering of the data, which isolates each worm-like data cloud and allows each object to be localized and tracked.

@@

A finite mixture model can be thought of as a probability distribution for an observation $x_{i}$ formulated as a linear combination of K mixture components (also called `clusters'), where the coefficients of the linear combination sum to one; each mixture component, in turn, is a probability distribution for $x_{i}$, given some parametric form. The finite mixture model can be written as
\begin{equation}
P(x) = \sum_{k=1}^{K} P(c_{i} = k)P(x|\theta_{k})
\end{equation}
where $c_{i} \in \{ 1, \ldots, K \}$ denotes the assignment of $x_{i}$ to a given mixture, which allows the coefficients of the linear combination to be written as $P(c_{i} = k)$. Additionally, $\theta_{k}$ denotes the parametric form of the $k^{\text{th}}$ mixture component. We also define $p_{k} = P(c_{i} = k)$ for all $k \in \{ 1, \ldots, K \} $. This model can also be written generatively as
\begin{align}
\begin{split}
	x_{i}|c_{i}, \theta_{c_{i}}  &\sim  F(\theta_{c_{i}}) \\
	c_{i}|p_{1}, \ldots, p_{K}  &\sim  \text{Discrete}(p_{1}, \ldots, p_{K}) \\
\end{split}
\end{align}
where the $x_{i}$ (for $i \in \{ 1, \ldots, N \}$) are observations, the $c_{i}$ are the mixture component assignments associated with each observation, and the $\theta_{c_{i}}$ are parameters defining the $c_{i}^{\text{th}}$ mixture component (i.e. the distribution to be mixed, $F(\theta_{c_{i}})$).

@@

The finite mixture model of the previous section can be extended to a Bayesian mixture model by viewing the distribution parameters that were previously point values, $\theta_{c_{i}}$ (the mixture component parameters) and $p_{1}, \ldots, p_{K}$ (the mixture component assignment weights), as random variables and providing each with a prior distribution. In this case, the prior distribution $\mathbb{G}_{0}$ is placed on the mixture component parameters, and the prior distribution $\text{Dir}(\alpha/K, \ldots, \alpha/K)$ is placed on the K mixture component assignment weights. The resulting Bayesian mixture model can be formulated in a hierarchical representation as
\begin{align}
\begin{split}
	\label{bayesian_mixture_model}
	x_{i}|c_{i}, \theta_{c_{i}}  &\sim  F(\theta_{c_{i}}) \\
	c_{i}|p_{1}, \ldots, p_{K}  &\sim  \text{Discrete}(p_{1}, \ldots, p_{K}) \\
	\theta_{c_{i}}  &\sim  \mathbb{G}_{0} \\
	p_{1}, \ldots, p_{K}  &\sim  \text{Dir}(\alpha/K, \ldots, \alpha/K)
\end{split}
\end{align}
where the $x_{i}$ (for $i \in \{ 1, \ldots, N \}$) are observations, the $c_{i}$ are the mixture component assignments associated with each observation, the $\theta_{c_{i}}$ are parameters defining the $c_{i}^{\text{th}}$ mixture component (i.e. the distribution to be mixed, $F(\theta_{c_{i}})$), the $\theta_{c_{i}}$ are drawn from the prior distribution $\mathbb{G}_{0}$, and $p_{1}, \ldots, p_{K}$ are drawn from a Dirichlet prior parameterized by $\alpha/K, \ldots, \alpha/K$.

@@

A Dirichlet process mixture model, also refered to as an infinite mixture model, is an extension of the Bayesian mixture model described previously. 
When using a DP as a prior in a Bayesian mixture model, $\Theta$ can be viewed as the set of parameters of the component mixture distributions. 	A DPM is formulated by allowing the prior distribution over component weights in a standard mixture model to be distributed according to a DP \willie{need to make sure am saying this right}; this allows for modeling data where the true number of latent mixture components is unknown and arbitrarily large by assuming an infinite number of components, of which only a finite amount are expressed by the data. In particular, the DPM can be defined generatively as
\begin{align}
\begin{split}
	x_{i}|\phi_{i} &\sim F(\phi_{i}) \\
	\phi_{i} | \mathbb{G}  &\sim  \mathbb{G} \\
	\mathbb{G} | \alpha, \mathbb{G}_{0}  &\sim  \text{DP}(\alpha, \mathbb{G}_{0})
\end{split}
\end{align}
where the $x_{i}$ (for $i \in \{ 1, \ldots, N \}$) are observations, the $\phi_{i}$ are parameters defining the mixture component from which the $i_{th}$ observation is drawn (i.e. the distribution to be mixed, $F(\phi_{i})$), and the $\phi_{i}$ are drawn from $\mathbb{G}$, which is in turn drawn from a DP with base distribution $\mathbb{G}_{0}$ and parameter $\alpha$. See \cite{gasthaus_2008} and \cite{gasthaus_thesis} for more details on this formulation. Note the difference between the indexing of the clusters in this model and the indexing in the previous two models. This formulation can be shown to be equivalent to the Bayesian mixture model defined in \eqref{bayesian_mixture_model}, when K is taken to be unbounded. This is the origin for references to this model as the infinite mixture model.

@@

\begin{figure}
  \centering               
  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/cross_final}}
  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/bounce_final}}
  \caption{Each plot shows a sample from the posterior distribution of the state for one of the videos in synthetic experiment one, where the time axis ascends vertically, the horizontal axes represent spatial position, color represents assignment, and the mean and standard deviation are shown. In both cases, the two objects are successfully tracked through occlusion, though the objects reverse direction in the second video. \willie{will show assignments by marker type in addition to color}}
  \label{fig:synth_one_plot}
\end{figure}

@@

Due to the computation required for the MCMC batch inference method (discussed further in Section~\ref{sec:discussion}), only the final 1000 frames of the video were used. Extraction was performed with frame differencing as described in Section~\ref{sec:modelspec_extraction}, using $L=3$. Parameter values were chosen in the same way, and set to the same values, as in the the synthetic experiments ($\alpha = 0.1, \rho = 0.3, M = 10, \boldsymbol{\mu}_{0} = (0,0), \kappa_{0} = 0.05, \nu_{0} = 5, \Lambda_{0} = \left( \begin{smallmatrix} 1&0\\ 0&1 \end{smallmatrix} \right), \text{ and } q_{0} = (5, \ldots, 5)$). The MCMC sampler was successful for both benchmark videos; each object was detected, tracked, and its shape estimated in manner very consistent with the ground-truth. The results for the PETS2000 dataset are displayed in Figure~\ref{fig:pets2000_results} and for the PETS2001 dataset in Figure~\ref{fig:pets2001_alphaimg}; in these figure, a sample from the posterior of the cluster parameters is overlayed the data and video frames at three time points, where the data is colored by assignment. \willie{also show pets 2001 result and per-frame overlays for both}.

@@

\begin{figure}[h]
  \centering             
  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2001_1}}
  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2001_2}}
  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2001_3}}
  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2001_4}}
  \caption{Four frames $f=25, 35, 45, 55$ from the PETS2001 sequence with one posterior sample mean and covarance matrix overlaid. \willie{will show path trail over prev 10 frames} \willie{these are old images!... haven't replaced yet}}
  \label{fig:pets2001_imgs}
\end{figure}

@@

Due to the large number of frames and objects in this video, the SMC algorithm (described in Section~\ref{sec:SMC}) was used for inference. This method of sequential inference was observed, on this dataset, to converge to a better sample in a shorter period of time in comparison with the MCMC algorithm. For this reason, we found that it was better suited for this large dataset.

>>>>

We define the task of automated detection and tracking of arbitrary objects in videos to be the fulfillment of three actions: determining the spatiotemporal regions of a video that constitute objects (`extraction'), finding the positions and/or shapes of distinct objects (`localization'), and maintaining the identities of the detected objects over time (`tracking'). An ability to carry out these tasks in an automated manner is useful for many fields that make use of video data, including robotics, video surveillance, time-lapse microscopy, and video summarization. Algorithms able to be applied to a variety of objects and video types are particularly desirable. This paper serves to introduce a new technique involving the use of dependent Dirichlet processes mixture models, which we hope will provide the foundation for a class of unsupervised algorithms that can detect and track arbitrary objects in a wide range of videos.

Research in the past ten years involving the development of methods for general detection and tracking of objects has tended to focus on one of either extraction, localization, or tracking; attempts to integrate all three tasks and produce a system aimed at multiple arbitrary objects and diverse video types was not often a primary focus, or was attempted in an ad-hoc manner to test or apply a more specific technique. Research aimed at developing methods to accomplish the three tasks in a cohesive manner has been pursued more recently to produce algorithms that can, in a fully automated way, accomplish unsupervised detection and tracking--or in short, that can determine the spatiotemporal region of a video occupied by each distinct object. A widely applied, though unsophisticated, example of general detection and tracking is the technique of blob tracking. Blob tracking typically refers to methods that perform extraction and localization of objects at each frame independently--for example, by assuming that pixels representing an object are distinguishable from pixels representing the background via color, motion, or texture information, and segmenting the extracted data at each frame into ``blobs'' to accomplish localization. A number of tracking techniques based on finding blobs in consecutive frames of a video with similar locations and appearances have been attempted to maintain the identity of detected blobs over time. The main challenges, however, associated with blob tracking involve the difficulty of segmenting neighboring or occluding objects in a given frame \cite{zhao2004tracking} and preventing multiple tracked blobs from collapsing onto and tracking a single object \cite{vermaak_2003}. These difficulties stem from the typical blob tracker's need to perform detection via segmentation on individual frames and lack of a principled way to perform tracking while resisting the collapse of independent tracks.

This paper presents the use of a Bayesian nonparametric mixture model for general, unsupervised multiple object detection and tracking. This method is intended to provide a widely applicable way to detect arbitrary objects in videos--particularly in cases where frame-by-frame segmentation is difficult or video quality is low and extraction is noisy--and better maintain the isolation of independent objects during tracking. Specifically, this work describes how a Dirichlet process mixture model can accomplish these goals when it is applied to data that may be extracted in a general manner from videos. We will demonstrate how to extract data from objects in videos and how this generative model allows for automated detection and tracking. The model is first presented in an abstract form, to allow it to be used with arbitrary extraction procedures and object appearance representations that may be desired in future studies. We then formulate a specific model for data gained during a basic extraction procedure and describe inference algorithms that allow object localization and tracking to be carried out. We demonstrate our implementation on a number of synthetic and benchmark datasets and describe and compute performance metrics to quantify results.

@@

Extraction is related to the tasks of foregound segmentation, background modeling, motion detection, and appearance feature extraction. In general, these methods aim to extract features from regions of a video frame that are considered objects or are distinguished in some way from the background of a scene. Areas of research in this field include attempts to handle situations where the background displays motion or other dynamic characteristics, reduce noise caused by the false extraction of backgrounds, handle abberant image effects present in some videos, and operate accurately on scenes involving illumination changes. These procedures can usually be classified as either model-based methods or heuristic-based methods. Model-based methods emphasize the detection of shapes of regions containing objects accurately and reducing background noise. For example, foreground extraction has been robustly carried out by modeling each pixel over time in order to infer a probability distribution over pixel values; if the pixel takes on a very unlikely value at a future time, it is marked as a foreground pixel at that point \citep{stauffer_1999, elgammal_2000, elgammal_2002}. Note that these techniques treat each pixel independently; this has a tendency to cause errors in foreground detection, resulting in fragmentation of foreground images (which is problematic as many of these methods intend to blob the foreground objects after detection in order to perform segmentation and locate objects' centroids). Heuristic-based methods perform extracting by looking for temporal changes in pixel value characteristics. These methods are often less accurate than model-based methods, but also far less computationally expensive. Frame differencing and background subtraction are two such methods, and involve comparing pixel values, or groups of pixel values, between two images, to detect a change surpassing some threshold. The goal of these techniques is to find locations in a scene that display motion, and to deem these moving areas the foreground of a scene.  Often, background subtraction refers to comparing an image containing targets with an image of the background without targets or with some model of the background that is learned as the video progresses (in these cases, objects that do not move throughout a scene will be treated as the background), while frame or image differencing refers to comparing two subsequent images in a video. Frame differencing has been used as the sole extraction method for object localization or tracking schemes with success \citep{pece_2002, beleznai_2006, chu_2007}, and also as a secondary data extraction method to help improve the accuracy of object tracking schemes \citep{perez_2002}. In addition to the two techniques mentioned above, there exist other ways of extracting movement from an image, such as through techniques involving optical flow (a calculation based on the `generalized gradient model' of an image, which attempts to capture the speed and direction of movement over areas in the image) \citep{horn_1981, bobick_2001}. In \citep{black_2000}, a motion detection and object tracking method is developed based on the relative motion of moving objects' boundaries, which are again found through analysis of optical flow.

Localization is closely related to the topics of object detection, position-finding, segmentation, and target recognition. These methods are mainly concerned with finding the positions and/or shapes of objects in a given video frame, or with discerning between objects similar in position or appearance. There has been a focus in this field on developing techniques for representing an object's appearance, both to allow for localization of objects on a per-frame basis and to allow for consistent localization of a given object over time. Research into the unsupervised detection of arbitrary objects (which we define as methods that accomplish both extraction and localization) has included work involving image segmentation, parsing, and pixel clustering \willie{must cite here}. These topics usually relate to the automated partitioning of images into meaningful regions. Note that the popular phrase `object detection' is also typically used to refer to procedures that carry out extraction and localization; however, it is often reserved to denote methods that are tailored to find the position of specified objects or object types (as opposed to arbitrary object detection), or to denote supervised methods, where extraction is peformed implicitely during searches for specified objects and localization procedures.

Tracking is related to the topics of video tracking, data association, filtering, point matching, and `tracking of non-rigid objects'. In general, these tasks aim to maintain the localization of the position and/or shape of an object over time. Areas of interest in this field involve developing metrics for judging the similarity between two object representations (for example, potential representations of the same object at adjacent frames in a video), techniques that provide a way to conduct efficient searches over a frame for regions that are similar to a given object, and methods for predicting the future position or appearance of an object.

@@

\subsection{Data}
\label{sec:data}
%
We defined the task of extraction to be the unsupervised determination of which spatiotemporal regions of a video constitute objects, and outlined a number of extraction methods in Section~\ref{sec:relatedwork}. Extraction allows us to generate a collection of data points. Each data point can be thought of as a hypothesis about, or noisy version of, the representation of an object; a collection of hypotheses for a given object is assumed to be distributed in some way around the true object representation. A video of multiple objects is therefore assumed to yield data that contains multiple groups of object hypotheses with similar characteristics (i.e. clusters of similar hypotheses).

Determining the regions of each video frame that contain objects allows us to generate these object hypotheses; namely, we let each hypothesis be a description of the pixel data over an area centered at a point in the extracted region. We can, for example, let each pixel in the extracted region be a hypothesis, and let the features of the representation consist of the spatial position of the pixel. An extraction dataset of this form is comprised of the three-dimensional points $\bold{x} = ( x_{1}, x_{2}, t ) \in \mathbb{R}^{2} \times \mathbb{Z}_{+}$, where $x_{1}$ and $x_{2}$ are the two spatial dimensions of a pixel contained in an extracted region in a given frame, and $t$ is the discrete time index of the frame. These three features are always available once regions containing objects are discerned, and are useful features for modeling (as position information is key for reasonable tracking and time information is necessary for maintaining the temporal dependencies in the data).

@@

We will use the phrase ``baseline'' three-dimensional extraction data to refer to the features $\bold{x} = ( x_{1}, x_{2}, t ) \in \mathbb{R}^{2} \times \mathbb{Z}_{+}$ of pixels present in regions containing objects. These data are distributed roughly as worm-like shapes that follow the paths of individual objects as they move, when they are plotted in three dimensions (where we let the frame number $t$, our discrete representation of time, ascend the vertical axis and the spatial positions occupy the horizontal axes). An example of data with such a distribution can been seen in Figure~\ref{fig:baselinedata}. The data consists of the three-dimensional baseline features and was gained by extraction peformed on a 50 frame sequence showing five ants moving across the video scene.

% \begin{figure}[h]
%   \centering
%   \subfloat[]{\label{fig:baselinedata1}\includegraphics[width=0.33\textwidth]{../img/pets2001_data_1.png}} 
%   \subfloat[]{\label{fig:baselinedata2}\includegraphics[width=0.33\textwidth]{../img/usbh_data_3.png}} 
%   \subfloat[]{\label{fig:baselinedata3}\includegraphics[width=0.33\textwidth]{../img/traffic2_data_1.png}}\\
%   \subfloat[]{\label{fig:baselinedata4}\includegraphics[width=0.33\textwidth]{../img/pets2009_data_1.png}} 
%   \subfloat[]{\label{fig:baselinedata5}\includegraphics[width=0.33\textwidth]{../img/pets2009_data_2.png}} 
%   \subfloat[]{\label{fig:baselinedata6}\includegraphics[width=0.33\textwidth]{../img/pets2009_data_3.png}}
%   \caption{Baseline extracted features from the (a.) PETS2001 Dataset, (b.) Ants Dataset, (c.) Traffic Dataset, and PETS2009/2010 benchmark dataset, frames (a.) 1-200, (b.) 201-400, (c.) 401-600. Time is shown on the vertical axis. Marker color corresponds with spatial coordinates and is shown for ease of viewing.}
%   \label{fig:baselinedata}
% \end{figure}

Inference performed on the time dependent Dirichlet process mixture model (Section~\ref{sec:inference}) carries out clustering of the data, which isolates each worm-like data cloud and allows each object to be localized and tracked.

@@

A finite mixture model can be thought of as a probability distribution for an observation $x_{i}$ formulated as a linear combination of K mixture components (also called `clusters'), where the coefficients of the linear combination sum to one; each mixture component, in turn, is a probability distribution for $x_{i}$, given some parametric form. The finite mixture model can be written as
\begin{equation}
P(x) = \sum_{k=1}^{K} P(c_{i} = k)P(x|\theta_{k})
\end{equation}
where $c_{i} \in \{ 1, \ldots, K \}$ denotes the assignment of $x_{i}$ to a given mixture (which allows the coefficients of the linear combination to be written as $P(c_{i} = k)$), and $\theta_{k}$ denotes the parametric form of the $k^{\text{th}}$ mixture component. We also define $p_{k} = P(c_{i} = k)$ for $k \in \{ 1, \ldots, K \} $. We can therefore write this model generatively as
\begin{align}
\begin{split}
	c_{i}|p_{1}, \ldots, p_{K}  &\sim  \text{Discrete}(p_{1}, \ldots, p_{K}) \\
	x_{i}|c_{i}, \theta_{c_{i}}  &\sim  F(\theta_{c_{i}})
\end{split}
\end{align}
where the $x_{i}$ (for $i \in \{ 1, \ldots, N \}$) are observations, the $c_{i}$ are the mixture component assignments associated with each observation, and the $\theta_{c_{i}}$ are parameters defining the $c_{i}^{\text{th}}$ mixture component (i.e. the distribution to be mixed, $F(\theta_{c_{i}})$).

@@

The finite mixture model of the previous section can be extended to a Bayesian mixture model by viewing the distribution parameters that were previously point values, $\theta_{c_{i}}$ (the mixture component parameters) and $p_{1}, \ldots, p_{K}$ (the mixture component assignment weights), as random variables and providing each with a prior distribution. In this case, the prior distribution $\mathbb{G}_{0}$ is placed on the mixture component parameters, and the prior distribution $\text{Dir}(\alpha/K, \ldots, \alpha/K)$ is placed on the K mixture component assignment weights. The resulting Bayesian mixture model can be formulated in a generative representation as
\begin{align}
\begin{split}
\label{bayesian_mixture_model}
	p_{1}, \ldots, p_{K}  &\sim  \text{Dir}(\alpha/K, \ldots, \alpha/K)\\
	c_{i}|p_{1}, \ldots, p_{K}  &\sim  \text{Discrete}(p_{1}, \ldots, p_{K}) \\
	\theta_{c_{i}}  &\sim  \mathbb{G}_{0} \\
	x_{i}|c_{i}, \theta_{c_{i}}  &\sim  F(\theta_{c_{i}})
\end{split}
\end{align}
where the $x_{i}$ (for $i \in \{ 1, \ldots, N \}$) are observations, the $c_{i}$ are the mixture component assignments associated with each observation, the $\theta_{c_{i}}$ are parameters defining the $c_{i}^{\text{th}}$ mixture component (i.e. the distribution to be mixed, $F(\theta_{c_{i}})$), the $\theta_{c_{i}}$ are drawn from the prior distribution $\mathbb{G}_{0}$, and $p_{1}, \ldots, p_{K}$ are drawn from a Dirichlet prior parameterized by $\alpha/K, \ldots, \alpha/K$.

@@

A Dirichlet process mixture model, also refered to as an infinite mixture model, is an extension of the Bayesian mixture model described previously. 
When using a DP as a prior in a Bayesian mixture model, $\Theta$ can be viewed as the set of parameters of the component mixture distributions. 	A DPM is formulated by allowing the prior distribution over component weights in a standard mixture model to be distributed according to a DP \willie{need to make sure am saying this right}; this allows for modeling data where the true number of latent mixture components is unknown and arbitrarily large by assuming an infinite number of components, of which only a finite amount are expressed by the data. In particular, the DPM can be defined generatively as
\begin{align}
\begin{split}
	\mathbb{G} | \alpha, \mathbb{G}_{0}  &\sim  \text{DP}(\alpha, \mathbb{G}_{0}) \\
	\phi_{i} | \mathbb{G}  &\sim  \mathbb{G} \\
	x_{i}|\phi_{i} &\sim F(\phi_{i})
\end{split}
\end{align}
where the $x_{i}$ (for $i \in \{ 1, \ldots, N \}$) are observations, the $\phi_{i}$ are parameters defining the mixture component from which the $i_{th}$ observation is drawn (i.e. the distribution to be mixed, $F(\phi_{i})$), and the $\phi_{i}$ are drawn from $\mathbb{G}$, which is in turn drawn from a DP with base distribution $\mathbb{G}_{0}$ and parameter $\alpha$. See \cite{gasthaus_2008} and \cite{gasthaus_thesis} for more details on this formulation. Note the difference between the indexing of the clusters in this model and the indexing in the previous two models. This formulation can be shown to be equivalent to the Bayesian mixture model defined in \eqref{bayesian_mixture_model}, when K is taken to be unbounded. This is the origin for references to this model as the infinite mixture model.

@@

\begin{figure}
  \centering               
  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/cross_final}}
  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/bounce_final}}
  \caption{Each plot shows a sample from the posterior distribution of the state for one of the videos in synthetic experiment one, where the time axis ascends vertically, the horizontal axes represent spatial position, color represents assignment, and the mean and standard deviation are shown. In both cases, the two objects are successfully tracked through occlusion, though both objects reverse their directions in the second video. \willie{will show assignments by marker type in addition to color}}
  \label{fig:synth_one_plot}
\end{figure}

@@

Due to the computation required for the MCMC batch inference method (discussed further in Section~\ref{sec:discussion}), only the final 1000 frames of the video were used. Extraction was performed with frame differencing as described in Section~\ref{sec:modelspec_extraction}, using $L=3$. Parameter values were chosen in the same way, and set to the same values, as in the the synthetic experiments ($\alpha = 0.1, \rho = 0.3, M = 10, \boldsymbol{\mu}_{0} = (0,0), \kappa_{0} = 0.05, \nu_{0} = 5, \Lambda_{0} = \left( \begin{smallmatrix} 1&0\\ 0&1 \end{smallmatrix} \right), \text{ and } q_{0} = (5, \ldots, 5)$). The MCMC sampler was successful for both benchmark videos; each object was detected, tracked, and its shape estimated in manner very consistent with the ground-truth. The results for the PETS2000 dataset are displayed in Figure~\ref{fig:pets2000_results} and for the PETS2001 dataset in Figure~\ref{fig:pets2001_alphaimg}; in these figure, a sample from the posterior of the cluster parameters is overlayed the data and video frames at three points in time, where the data is colored by assignment. \willie{also show pets 2001 result and per-frame overlays for both}.

@@

\begin{figure}[h]
  \centering             
  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2001_1}}
  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2001_2}}
  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2001_3}}
  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2001_4}}
  \caption{Four frames $f=25, 35, 45, 55$ from the PETS2001 sequence with one posterior sample mean and covarance matrix overlaid. \willie{will show path trail over prev 10 frames.} \willie{These are old images!... haven't replaced yet}}
  \label{fig:pets2001_imgs}
\end{figure}

@@

Due to the large number of frames and objects in this video, the SMC algorithm (described in Section~\ref{sec:SMC}) was used for inference. This method of sequential inference was observed, on this dataset, to converge to a better sample in a shorter period of time in comparison with the MCMC algorithm; for this reason, we found that it was better suited for this large dataset.

Extraction was performed with frame differencing as described in Section~\ref{sec:modelspec_extraction}, using $L=3$ and parameters for the model were chosen, using the method outlined in Section~\ref{sec:parameter_recap}, to be $\alpha = 0.1, \rho = 0.8, M = 10, \boldsymbol{\mu}_{0} = (0,0), \kappa_{0} = 0.05, \nu_{0} = 6, \Lambda_{0} = \left( \begin{smallmatrix} 1&0\\ 0&1 \end{smallmatrix} \right), \text{ and } q_{0} = (3, \ldots, 3)$. Additionally, as with the other video datasets, ground-truth bounding boxes around each object were authored using the ViPER tool.

The SMC inference algorithm yielded a sample from the posterior distribution, from which the object detection and tracking results were obtained (as described in Section~\ref{sec:inference_to_results}). Figure \willie{provide figure showing PETS2009 results} shows a sample from the posterior of the cluster parameters overlayed on the data, and video frames at three points in time, where the data is colored by assignment
\willie{show SFDA and ATA vs cov mat conf for this sample that I show figures for.}
